\documentclass[twoside,11pt]{article}
\usepackage{jagi}
\newcommand{\dataset}{{\cal D}}
\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
\jagiheading{0}{2008}{1-?}{2008-??-??}{2008-??-??}{Moshe Looks}
\ShortHeadings{Heuristic Bayesian Program Evolution}{Looks}
\firstpageno{1}

\begin{document} \sloppy
\title{Heuristic Bayesian Program Evolution}
\author{\name Moshe Looks \email madscience@google.com \\
       \addr Google, Inc. \\
       1600 Amphitheatre Pkwy \\
       Mountain View, CA 94043, USA}
\editor{Pei Wang}
\maketitle

\begin{abstract}
  Solomonoff's formalization of inductive inference and Wang's assumption of
  insufficient knowledge and resources provide the basic theoretical guidance
  and constraints necessary for the design of an artificial general
  intelligence, implementable on near-future hardware. I further hypothesize
  that approximations for computationally intractable feats of Bayesian
  inductive inference may be obtained through judiciously assimilated
  background knowledge. This paper has a tripartite structure. Section one
  describes and argues my premises in greater detail. Section two sketches the
  design of a system embodying these principles that is under development, and
  presents some results. Section three discusses appropriate goals for AGI
  systems to ensure benevolence towards humans, and speculates how such goals
  might be implemented.
\end{abstract}

\begin{keywords}
  program evolution, Bayes, inductive inference, friendliness
\end{keywords}

\section{Normal Forms}

foo.

\acks{goo.}

\bibliography{refs}
\end{document}
